---
title: "Time Series Forecast of Airline Ticket Sales"
output:
  html_document:
    df_print: paged
---
## Objective: 
To predict the airline tickets' sales of 1961 using Time Series Analysis.

## Data Description:
10 year air-ticket sales data of airline industry from 1949-1960



```{r}
#Install and load parkages
#install.packages("forecast")
#install.packages("tseries")
```

```{r}
library(forecast)
library(tseries)
```

#### Exploratory Data Analysis
```{r}
data("AirPassengers")
class(AirPassengers)
```

```{r}
start(AirPassengers)
end(AirPassengers)
frequency(AirPassengers)
```

```{r}
sum(is.na(AirPassengers))

```

```{r}
summary(AirPassengers)
AirPassengers
```



### INSIGHTS
1. **Trend**: The passenger numbers increase over time indicating an increasing linear trend.
2. **Seasonality**: In the boxplot, there are more passengers travelling in months 6 to 9, indicating seasonality with an apparent cycle of 12 months.

## Decomposing Time Series into its components

```{r}
tsdata<-ts(AirPassengers, frequency = 12)
ddata <- decompose(tsdata, "multiplicative")
```

```{r}
plot(ddata)

```

#### trend component
```{r}
plot(ddata$trend)
```

#### seasonality component
```{r}
plot(ddata$seasonal)

```

#### Irregularity component
```{r}
plot(ddata$random)
```

#### Plotting components together

```{r}
plot(AirPassengers)
abline(reg = lm(AirPassengers~time(AirPassengers)))
cycle(AirPassengers)
```

#### Boxplot by cycle
```{r}
boxplot(AirPassengers~cycle(AirPassengers, xlab="Date"))

```

#### Test for Stationarity
```{r}
plot(AirPassengers)

```

#### Find the best ARIMA Model
```{r}
mymodel <- auto.arima(AirPassengers)
mymodel
```


#### Run best model with `trace` to compare the information criteria
```{r}
auto.arima(AirPassengers, ic="aic", trace=TRUE)

```

-- #adf.test(mymodel)

## Checking residuals which shows stationarity
```{r}
plot.ts(mymodel$residuals)

```

#### Plot Autocorrelation Function (ACF) and Partial ACF
```{r}
acf(ts(mymodel$residuals), main="ACF Residual")
pacf(ts(mymodel$residuals), main="PACF Residual")

```

### Forecast for the next 10 years
```{r}
myforecast<-forecast(mymodel, level=c(95), h=10*12)
plot(myforecast)
```


### Validating the findings of the ARIMA model 
```{r}
Box.test(mymodel$resid, lag=5, type="Ljung-Box")
Box.test(mymodel$resid, lag=10, type="Ljung-Box")
Box.test(mymodel$resid, lag=15, type="Ljung-Box")
```


### Results
The p-values are quite insignificant, indicating that the model is free of autocorrelation.

### Conclusion
From the ARIMA output, the model using (2,1,1) has been shown to adequately fit the data. 



